---
title: "Lullington Heath: a Long-Term Monitoring Network report"
#author: "Kieran Fox"
#date: "22/02/2021"
#output: html_document
output:
  html_document:
    css: "./styles/air.css"
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, echo=FALSE, include=FALSE}
source('./data_prep_figure4.R')
knitr::opts_chunk$set(echo = FALSE)
opts_chunk$set(fig.width=8.65, fig.height=5)
setwd('C:/Users/kiera/Work/NE_work/Report')
```

```{r import, include=FALSE}
name_swap_bap <- get_names('./dataframes/rename_habitat.csv')[[1]]
name_swap_nvc <- get_names('./dataframes/rename_habitat.csv')[[2]]
bb_list <- get_names('./dataframes/rename_habitat.csv')[[3]]
bp_list <- get_names('./dataframes/rename_habitat.csv')[[4]]

#df_all <- read_all_data('all_plots.csv')
df_all <- read_csv('./dataframes/report_plots.csv')
site_code <- 'B14'

# Taking the data from just one site
df_site <- filter(df_all, Sitecode == site_code) %>%
  fix_names(., name_swap_bap) %>%
  # transform the eastings and northings into latitude and longitude
  transform_coords(.) %>%
  distinct(Plot_ID, Sitecode, Year, .keep_all = TRUE)


# Converts the NVC abbreviations to the full names
df_site$NVC_habitat <- name_swap(df_site$NVC_habitat, name_swap_nvc)
df_site$Year <- as.character(df_site$Year)

df_site <- rem_plot(df_site, c(30, '30', '30a')) # Lullington
#df_site <- rem_plot(df_site, c(14, '14', '14a')) # thursley

# Making the sub dataframes with only the major habitats selected
#object[[1]] is the data frame. object[[2]] is the list of habitats
df_site.bb <- select_by(df_site, 'BAP_broad', 700, include = c('Calcareous Grassland', 'Dwarf Shrub Heath'))
df_site.bp <- select_by(df_site, 'BAP_priority', 15)
df_site.nvcb <- select_by(df_site, 'NVC_group', 15)
df_site.nvcc <- select_by(df_site, 'NVC_habitat', 500, include = c('Calcareous grassland (CG)', 'Heath (H)'))

# the centre points of the maps
east_cent <- get_centre_coords(df_site)[[1]]
north_cent <- get_centre_coords(df_site)[[2]]

unique_years <- unique(df_site$Year)
unique_habs <- unique(df_site$BAP_broad)
unique_nvchabs <- unique(df_site$NVC_habitat)

df_change <- df_site %>%
  get_change_by_year() %>%
  distinct(Plot_ID, Year, .keep_all = TRUE)

hab_sum <- read_csv('./dataframes/all_plots.csv') %>%
  get_hab_sums()

#df_change <- rem_plot(df_change, c(14, '14', '14a')) # thursley
```

## 

The information presented here is a summary of the vegetation data collected throughout the LTMN project. The ultimate aim of LTMN is to survey over long periods of time in order to distinguish between trends and the fluctuations inherent in the environment. We are starting to reach three surveys for each site so we can start to see trends but as yet, they are difficult to have confidence in due to the limited number of data points. However, we want to make the data available so that any information that can be gleaned is accessible. This form of report has been structured so that as each new survey is implemented, the data is added to this report and will be available rapidly after data collection. The content is also being updated and we would greatly appreciate any feedback on the kind of information that is most useful, which parts of the report are difficult to use and any site specific data to present (such as individual species or groups of species). 

We would like to introduce the caveat that while we have attempted to conduct the survey at a similar date in the year, there can be some small changes between surveys due to logistical difficulties. This might have affected the species present.

```{r dates_of_survey}
display_survey_dates(df_site)
```

## Habitats {.tabset .tabset-fade}

The sites of the LTMN project have been chosen to represent key Biodiversity Action Plan (BAP) habitats. Individual plots are assigned a BAP broad and priority habitat when data is collected, and where this categorisation is missing, it is filled using the other available data (such as NVC habitat classification). 

Currently we do not present BAP priority habitats as there is too much missing data, but the broad habitat gives a good indication of what the environment at the plot level looks like to the surveyor.

### BAP broad habitats

```{r Bap_habitat}
map_hab_by_year(df_site, 'BAP_broad', 'BAP_priority')#df_site.bb[[1]]
```

Each circle on the map represents a plot or quadrat. The locations of the plots are recorded so that they can be reused for each survey. This means plots can be compared from year to year and changes are likely due to changing habitat. The exact location is recorded using feno markers placed in the ground. Sometimes these are lost between surveys and in such cases, the coordinates are used to roughly match the location. This can mean that sometimes the plots move a couple of metres. In some cases, not all plots are surveyed each time, typically due to time restraints on the survey.

```{r number_of_habs_broad}
display_no_plots(df_site, 'BAP_broad', 'Species_richness', 
  'Number of each BAP broad habitat')
```

There are usually a mix of habitats on LTMN sites with certain habitats being more prominent. We take roughly 50 quadrats each survey which means that the lesser represented habitats usually make up around 2-4 quadrats each. We will in general omit these from the analysis as they are not numerous enough to trust that any changes are not due to chance.

### NVC habitats

After the species data has been collected for each plot, it is processed using MAVIS software which uses the plants and plant abundance to categorise the plots into their NVC habitat classifications. It should be noted that accurate categorisation is often difficult due to sites having a mix of habitats and transitions between them. For this reason the categorisation is given a confidence percentage which is shown when hovering the mouse over the data points on the map below.

```{r NVC_habitats}
#map_hab_by_year(df_site.nvcc[[1]], 'NVC_habitat', 'NVC_FIRST')
map_hab_by_year(df_site, 'NVC_habitat', 'NVC_FIRST')
```

Each circle on the map represents a plot or quadrat. The locations of the plots are recorded so that they can be reused for each survey. This means plots can be compared from year to year and changes are likely due to changing habitat. The exact location is recorded using feno markers placed in the ground. Sometimes these are lost between surveys and in such cases, the coordinates are used to roughly match the location. This can mean that sometimes the plots move a couple of metres. In some cases, not all plots are surveyed each time, typically due to time restraints on the survey.

```{r number_of_habs_nvc}
display_no_plots(df_site, 'NVC_habitat', 'Species_richness', 
  'Number of each NVC habitat group')
```

There are usually a mix of habitats on LTMN sites with certain habitats being more prominent. We take roughly 50 quadrats each survey which means that the lesser represented habitats usually make up around 2-4 quadrats each. We will in general omit these from the analysis as they are not numerous enough to trust that any changes are not due to chance.

### NVC habitat communities

Within each larger NVC group of communities (i.e. H, CG, W), there are many differences between the communities (i.e. CG2, CG6). For each major NVC group, the individual community split is displayed below over the years.

```{r NVC_subgroups}
habitat_by_habitat('NVC_habitat', 'NVC_group')
```

## Species {.tabset .tabset-fade}

```{r div_rich_map}
int_map_feat_by_hab(df_site, c('Species_richness', 'Species_diversity'))
```

Species richness (the number of different species found in each plot) and species diversity (a measure of how even the species populations within a plot are) are recorded and mapped onto the site. A darker shade of blue indicates a higher richness/diversity. Only plots from the most recent survey are included on this map.

### Species richness

Total number of unique species found in each plot is recorded. The distributions of species numbers in each habitat over multiple surveys can then be compared

##### Average change over the years

```{r av_change_spec_rich}
plot_feat_by_hab_year('BAP_broad', 'Species_richness')
```

The centre line of the box is the median value across all plots designated as that habitat. The edges of the box are the interquartile range (the range, omitting the smallest and largest quarters of the values).

```{r no_plots_spec_rich}
#display_no_plots(df_site.bb[[1]], 'BAP_broad', 'Species_richness')
```

```{r tot_av_spec_rich}
display_averages(hab_sum, 'Species_richness')
```

##### Map of change in species richness

```{r map_change_spec_rich}
change_map_int(df_change, 'Species_richness', norm=TRUE, leg_title='Percentage change')
```

Map showing where changes are located. Each year shows the change from the last survey across each plot, with purple markers indicating a reduction and yellow markers indicating an increase. The change in each plot has been normalised by dividing the change in species richness by the total plot species richness to give a percentage change.

### Species diversity

A community dominated by a small portion of the species is considered to be less diverse than one in which more of the species present have a similar abundance. The Simpson diversity index takes this into account as well as total species present. Values of diversity range from 0 to 1 with 0 meaning a total lack of biodiversity (monoculture) and 1 meaning a completely even spread of species.

##### Average change over the years

```{r av_change_spec_div}
plot_feat_by_hab_year('BAP_broad', 'Species_diversity')
```

The distribution of the plot diversity scores across the major BAP broad habitats. The centre line of the box is the median value across all plots designated as that habitat. The edges of the box are the interquartile range (the range, omitting the smallest and largest quarters of the values).

```{r no_plots_spec_div}
#display_no_plots(df_site.bb[[1]], 'BAP_broad', 'Species_richness')
```

```{r tot_av_spec_div}
display_averages(hab_sum, 'Species_diversity')
```

##### Map of change in species diversity

```{r map_change_spec_div}
change_map_int(df_change, 'Species_diversity')
```

Map showing where changes are located. Each year shows the change from the last survey across each plot, with purple markers indicating a reduction and yellow markers indicating an increase. 

### Individual species plot

```{r individual_species, echo=FALSE}
plot_ind_species('./dataframes/all_species_pc.csv', 
                 './dataframes/Positive_indicators.csv', 
                 site_code)
```

Hover the mouse over each line to get details. Click on species in the legend to add and remove them from the view.

## Vegetation structure {.tabset .tabset-fade}

```{r physical_map}
#df_site2 <- rem_plot(df_site, c(118, '118')) # martin down
#df_site2 <- rem_plot(df_site, c(14, '14', '14a')) # thursley
int_map_feat_by_hab(df_site, c('Vegetation_height', 'Litter', 'Bare_ground'))
```

This map shows a darker shade of blue for more litter/bare ground and higher plant height. Only plots from the most recent survey are included on this map. Vegetation height is an average across the whole plot. Litter and Bare ground is measured out of 25 and is the recording of presence or absence in each of the 25 quadrat subdivisions.

### Vegetation height

The vegetation height in a plot is measured by placing a disk ontop of the sward and letting it fall until it is supported by the vegetation. This is done in each division of the larger quadrat and the height is recorded. The average of these is then taken for each plot and is given in cm.

##### Average change over the years

```{r av_change_veg_height}
plot_feat_by_hab_year('BAP_broad', 'Vegetation_height')
```

The distribution of the plot vegetation height averages across the major BAP broad habitats. The centre line of the box is the median value across all plots designated as that habitat. The edges of the box are the interquartile range (the range, omitting the smallest and largest quarters of the values).

```{r no_plots_veg_height}
#display_no_plots(df_site.bb[[1]], 'BAP_broad', 'Species_richness')
```

```{r tot_av_veg_height}
display_averages(hab_sum, 'Vegetation_height')
```

##### Map of change in vegetatation height

```{r map_change_veg_height}
#df_change2 <- rem_plot(df_change, c(14, '14', '14a')) # thursley
change_map_int(df_change, 'Vegetation_height')
```

Map showing where changes are located. Each year shows the change from the last survey across each plot, with purple markers indicating a reduction and yellow markers indicating an increase. 

### Litter

Presence of litter is recorded in each of the 25 subdivisions of the larger quadrat and each plot is then given a recording of 0-25 for presence of litter.

##### Average change over the years

```{r av_change_Litter}
plot_feat_by_hab_year('BAP_broad', 'Litter')
```

The distribution of the litter presence scores across the major BAP broad habitats. The centre line of the box is the median value across all plots designated as that habitat. The edges of the box are the interquartile range (the range, omitting the smallest and largest quarters of the values).

```{r no_plots_litter}
#display_no_plots(df_site.bb[[1]], 'BAP_broad', 'Species_richness')
```

```{r tot_av_Litter}
display_averages(hab_sum, 'Litter')
```

##### Map of change in presence of litter

```{r map_change_Litter}
change_map_int(df_change, 'Litter')
```

Map showing where changes are located. Each year shows the change from the last survey across each plot, with purple markers indicating a reduction and yellow markers indicating an increase. 

### Bare ground

Presence of bare ground is recorded in each of the 25 subdivisions of the larger quadrat and each plot is then given a recording of 0-25 for presence of litter. 

##### Average change over the years

```{r av_change_bare}
plot_feat_by_hab_year('BAP_broad', 'Bare_ground')
```

The distribution of the bare ground presence scores across the major BAP broad habitats. The centre line of the box is the median value across all plots designated as that habitat. The edges of the box are the interquartile range (the range, omitting the smallest and largest quarters of the values).

```{r no_plots_bare}
#display_no_plots(df_site.bb[[1]], 'BAP_broad', 'Species_richness')
```

```{r tot_av_bare}
display_averages(hab_sum, 'Bare_ground')
```

##### Map of change in presence of bare ground

```{r map_change_bare}
change_map_int(df_change, 'Bare_ground')
```

Map showing where changes are located. Each year shows the change from the last survey across each plot, with purple markers indicating a reduction and yellow markers indicating an increase. 

## Ellenberg scores {.tabset .tabset-fade}

These are scores of 'light', 'fertility', 'pH', and 'wetness' for each plot, given by the plant species found. They are not a direct measurement of the feature but instead, plants are given a score for each, depending on how they like the particular feature. The plants in the plot are then averaged to give a plant favourability score for each plot. A high wetness score means the plants in that plot favour a wet environment and a low fertility score means the plants favour an relatively infertile soil.

Ellenberg scores where first assigned by Ellenberg _et al._ (1991) for a large number of european flora species, based on experimental work as well as field observation. These scores were assigned using a central European environment and so were later adapted by Hill _et al._ (1999, 2000) for the British environment. 

```{r ellenberg_map}
int_map_feat_by_hab(df_site, c('Light', 'Fertility', 'pH', 'Wetness'))
```

This map shows a darker shade of blue for higher Ellenberg scores. Only plots from the most recent survey are included on this map.

The pH scores here are not 'pH' but a value of alkaline favourability of the plants found on the plot. They have no units and do not scale in the same way that pH scales (with 7 as neutral). Large scores mean that the plants on the plot prefer alkaline soils.

### Light

The light Ellenberg score gives an indication of how the plants on a plot favour light. A score of 1 indicates the plant prefers deep shade at all times, a score of 5 indicates semi shade and a score of 9 indicates the plant favours full light.

##### Average change over the years

```{r av_change_light}
plot_feat_by_hab_year('BAP_broad', 'Light')
```

The distribution of the 'light' values across the major BAP broad habitats. The centre line of the box is the median value across all plots designated as that habitat. The edges of the box are the interquartile range (the range, omitting the smallest and largest quarters of the values). Plants are scored on a scale of 1-9 and this average is shown on the y-axis.

```{r no_plots_light}
#display_no_plots(df_site.bb[[1]], 'BAP_broad', 'Species_richness')
```

```{r tot_av_light}
display_averages(hab_sum, 'Light')
```

##### Map of change in light scores

```{r map_change_light}
change_map_int(df_change, 'Light')
```

Map showing where changes are located. Each year shows the change from the last survey across each plot, with purple markers indicating a reduction and yellow markers indicating an increase. 

### Fertility

The fertility Ellenberg score gives an indication of how the plants on a plot favour nitrogen in the soil. A score of 1 indicates the plant prefers minimal levels of nitrogen, a score of 5 indicates an average favourability and a score of 9 indicates the plant favours excessive supply.

##### Average change over the years

```{r av_change_fertility}
plot_feat_by_hab_year('BAP_broad', 'Fertility')
```

The distribution of the 'fertility' values across the major BAP broad habitats. The centre line of the box is the median value across all plots designated as that habitat. The edges of the box are the interquartile range (the range, omitting the smallest and largest quarters of the values). Plants are scored on a scale of 1-9 and this average is shown on the y-axis.

```{r no_plots_fertility}
#display_no_plots(df_site.bb[[1]], 'BAP_broad', 'Species_richness')
```

```{r tot_av_fertility}
display_averages(hab_sum, 'Fertility')
```

##### Map of change in fertility scores

```{r map_change_fertility}
change_map_int(df_change, 'Fertility')
```

Map showing where changes are located. Each year shows the change from the last survey across each plot, with purple markers indicating a reduction and yellow markers indicating an increase. 

### pH

A score of 1 indicates the plant prefers extremely acidic soils, a score of 5 means mildly acidic conditions are preferred, and a score of 9 indicates the plant favours alkaline soils.

##### Average change over the years

```{r av_change_ph}
plot_feat_by_hab_year('BAP_broad', 'pH')
```

The distribution of the 'pH' values across the major BAP broad habitats. The centre line of the box is the median value across all plots designated as that habitat. The edges of the box are the interquartile range (the range, omitting the smallest and largest quarters of the values). Plants are scored on a scale of 1-9 and this average is shown on the y-axis.

```{r no_plots_ph}
#display_no_plots(df_site.bb[[1]], 'BAP_broad', 'Species_richness')
```

```{r tot_av_ph}
display_averages(hab_sum, 'pH')
```

##### Map of change in pH scores

```{r map_change_ph}
change_map_int(df_change, 'pH')
```

Map showing where changes are located. Each year shows the change from the last survey across each plot, with purple markers indicating a reduction and yellow markers indicating an increase. 

### Wetness

The wetness Ellenberg score gives an indication of how the plants in the plot prefer the wetness of the soil. A score of 1 indicates dry soil is preferred, 5 would indicate the plants favour a moist soil, 9 suggest wet soil is preferred and a score of 10-12 indicates the plant favours underwater conditions.

##### Average change over the years

```{r av_change_wetness}
plot_feat_by_hab_year('BAP_broad', 'Wetness')
```

The distribution of the 'wetness' values across the major BAP broad habitats. The centre line of the box is the median value across all plots designated as that habitat. The edges of the box are the interquartile range (the range, omitting the smallest and largest quarters of the values). Plants are scored on a scale of 1-12 and this average is shown on the y-axis.

```{r no_plots_wetness}
#display_no_plots(df_site.bb[[1]], 'BAP_broad', 'Species_richness')
```

```{r tot_av_wetness}
display_averages(hab_sum, 'Wetness')
```

##### Map of change in wetness scores

```{r map_change_wetness}
change_map_int(df_change, 'Wetness')
```

Map showing where changes are located. Each year shows the change from the last survey across each plot, with purple markers indicating a reduction and yellow markers indicating an increase. 

## Grime's scores {.tabset .tabset-fade}

Universal adaptive strategy theory (UAST) is a method of categorising plants according to their survival strategy. There are three main ways in which a plant can specialise in order to take advantage of its environment. The first is by taking advantage of available of resources and maximising acquisition at the expense of competitors. Plants are given a 'competition' score to represent the extent to which they use this strategy. The second is by being able to deal with a small amount of resources while other species are not able to maintain in such a stressful environment. Plants are given a 'stress' score to represent the extent to which they use this strategy. The third and final strategy is with rapid completion of the life-cycle which can be advantageous in environments where events are frequently lethal to the plant. Plants are given a 'ruderals' score to represent the extent to which they use this strategy.

A large number of British species were placed in this triangle with assigned scores for stress, competition and ruderals by Grime _et al._ (1988, 1995)

```{r grimes_map}
int_map_feat_by_hab(df_site, c('Stress', 'Competition', 'Ruderals'))
```

Individual plots have been taken and the average 'stress', 'competition' and 'ruderals' scores are calculated. This map shows a darker shade of blue for higher Grime's scores. Only plots from the most recent survey are included on this map. 

### Competition

Plants with a high competition score thrive in environments with low disturbance and low intensity stress. They are able to outcompete other plants for the most valuable resources through rapid growth, physical spread and a high capacity for phenotype plasticity (enabling them to adapt to seasons and changing availablity of resources).

##### Average change over the years

```{r av_change_comp}
plot_feat_by_hab_year('BAP_broad', 'Competition')
```

The distribution of the 'competition' values across the major BAP broad habitats. The centre line of the box is the median value across all plots designated as that habitat. The edges of the box are the interquartile range (the range, omitting the smallest and largest quarters of the values). Plants are scored on a scale of 1-5 and this average is shown on the y-axis.

```{r no_plots_comp}
#display_no_plots(df_site.bb[[1]], 'BAP_broad', 'Species_richness')
```

```{r tot_av_comp}
display_averages(hab_sum, 'Competition')
```

##### Map of change in competition scores

```{r map_change_comp}
change_map_int(df_change, 'Competition')
```

Map showing where changes are located. Each year shows the change from the last survey across each plot, with purple markers indicating a reduction and yellow markers indicating an increase. 

### Stress

Plants with a high stress score have adapted to live in highly stressful environments such as alpine or arid habitats, deep shade, nutrient poor soils or areas with extreme pH levels. They are slow growing plants with high nutrient retention and do not do well with lots of disturbance.

##### Average change over the years

```{r av_change_stress}
plot_feat_by_hab_year('BAP_broad', 'Stress')
```

The distribution of the 'stress' values across the major BAP broad habitats. The centre line of the box is the median value across all plots designated as that habitat. The edges of the box are the interquartile range (the range, omitting the smallest and largest quarters of the values). Plants are scored on a scale of 1-5 and this average is shown on the y-axis.

```{r no_plots_stress}
#display_no_plots(df_site.bb[[1]], 'BAP_broad', 'Species_richness')
```

```{r tot_av_stress}
display_averages(hab_sum, 'Stress')
```

##### Map of change in stress scores

```{r map_change_stress}
change_map_int(df_change, 'Stress')
```

Map showing where changes are located. Each year shows the change from the last survey across each plot, with purple markers indicating a reduction and yellow markers indicating an increase. 

### Ruderals

Ruderals thrive in low stress environments but with lots of disturbance. They will quickly complete their lifecycle and often release are large amount of seed. Plants that have adapted to this strategy are often found in recently disturbed land and are often annuals.

##### Average change over the years

```{r av_change_ruderals}
plot_feat_by_hab_year('BAP_broad', 'Ruderals')
```

The distribution of the 'ruderals' values across the major BAP broad habitats. The centre line of the box is the median value across all plots designated as that habitat. The edges of the box are the interquartile range (the range, omitting the smallest and largest quarters of the values). Plants are scored on a scale of 1-5 and this average is shown on the y-axis.

```{r no_plots_ruderals}
#display_no_plots(df_site.bb[[1]], 'BAP_broad', 'Species_richness')
```

```{r tot_av_ruderals}
display_averages(hab_sum, 'Ruderals')
```

##### Map of change in ruderal scores

```{r map_change_ruderals}
change_map_int(df_change, 'Ruderals')
```

Map showing where changes are located. Each year shows the change from the last survey across each plot, with purple markers indicating a reduction and yellow markers indicating an increase. 

## 

For more information contact the LTMN team:

Project manager - kate.fagan@naturalengland.org.uk

Vegetation lead - sarah.grinstead@naturalengland.org.uk

Author and analyst - kieran.fox@naturalengland.org.uk

We are happy to provide explanations as well as further analysis.

### Bibliography

Ellenberg, H., Weber, H.E., Dull, R., Wirth, V., Werner, W., Paulissen, D. (1991). Zeigerwerte von Pflanzen in Mitteleuropa. Scripta Geobotanica. 18, 1-248

Hill, M.O., Mountford, J.O., Roy, D.B., Bunce, R.G.H. (1999) Ellenberg's indicator values for British plants: Ecofact Volume 2, Technical Annex. ITE Monkswood & ITE Grange-over-Sands. Centre for Ecology & Hydrology and Natural Environment Research Council

Hill M.O., Roy D.B., Mountford J.O. & Bunce R.G.H. 2000. Extending Ellenberg’s indicator values to a new area: an algorithmic approach. J. Appl. Ecol. 37: 3-15

Comparative Plant Ecology: A Functional Approach to Common British Species, Authors: Grime, J.P., Hodgson, J.G., Hunt, R.

Grime, J.P., Hodgson, J.G., Hunt, R. & Thompson, K. (1995). The Electronic Comparative Plant Ecology. London. Chapman & Hall